---
title: "Untitled"
author: "Kunal"
date: "5/14/2020"
output: html_document
---
 1. Load the data. 
 2. Process/transform the data.
 3. Calculate the total number of steps taken per day.
 4. Histogram of the total number of steps taken each day.
 
```{r , echo=TRUE}
activity<-read.csv("activity.csv")
data <- aggregate(steps ~ date, data=activity, sum, na.rm = TRUE)
hist(data$steps,breaks=20)
steps_mean <- mean(data$steps, na.rm=TRUE)
steps_median <- median(data$steps, na.rm=TRUE)
print(paste("The mean steps per day is: ", steps_mean))
print(paste("The median steps per day is: ", steps_median))

```
 
 1. Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).
 2. 5-minute interval, on average across all the days in the dataset, containing the maximum number of steps.

```{r echo=TRUE}
data2<-aggregate(steps ~ interval, data=activity, mean, na.rm=TRUE)
plot(data2$interval, data2$steps, type="l")
max_steps<-max(data2[,2])
print(paste("The maximum number of steps in a five minute interval is: ", max_steps))
```

1.The total number of missing values in the dataset 
2.Strategy for filling in all of the missing values in the dataset
3.A new dataset that is equal to the original dataset but with the missing data filled in.
4.Histogram of the total number of steps taken each day along with mean and median total number of steps taken per day.

```{r, echo=TRUE}
  total_na<-sum(is.na(activity))
print(paste("There are",total_na , "missing data points."))
betterdata <- activity
betterdata$steps[is.na(betterdata$steps)] <- median(activity$steps, na.rm=TRUE)
betterdataday <- aggregate(steps ~ date, data=betterdata, sum, na.rm=TRUE)
hist(betterdataday$steps, breaks=20)
bsteps_mean <- mean(betterdata$steps)
bsteps_median <- median(betterdata$steps)
print(paste("The mean is: ", bsteps_mean))
print(paste("The median is: ", bsteps_median))

```

1.Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
2.A panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r, echo=TRUE}
betterdata$date <- as.Date(betterdata$date)
betterdata$dayname <- weekdays(betterdata$date)
betterdata$weekend <- as.factor(ifelse(betterdata$dayname == "Saturday" |
                                 betterdata$dayname == "Sunday", "weekend", "weekday"))
plotdata <- aggregate(steps ~ interval + weekend, betterdata, mean)
library(lattice)
xyplot(steps ~ interval | factor(weekend), data=plotdata, aspect=1/3, type="l")














```

